---
title: "Week 6"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown



```{r message=FALSE, warning=FALSE}
library(nycflights13)
library(tidyverse)
library(Lahman)
```


###Look at the data to begin with
```{r message=FALSE, warning=FALSE}
flights
```

###filter the data for the 1st of january using filter() function
```{r message=FALSE, warning=FALSE}
filter(flights, month == 1, day == 1)
jan1 <- filter(flights, month == 1, day == 1)
```

###in case we need to both print the results and save it a new variable, then we need to enclose the data assignment in parantheses
```{r message=FALSE, warning=FALSE}
(dec25 <- filter(flights, month == 12, day == 25))
```


###Due to finite percision arithmetic, we cannot use == when comparing calculations the way we normally would. Eg: sqrt(2) ^ 2 == 2 would throw out FALSE
```{r message=FALSE, warning=FALSE}
near(sqrt(2) ^ 2,  2)
near(1 / 49 * 49, 1)
```

###Filter all flights from month = November or December
```{r message=FALSE, warning=FALSE}
filter(flights, month == 11 | month == 12)
```

###If we have multiple months, it becomes difficult to type this out, we can use %in% instead
```{r message=FALSE, warning=FALSE}
nov_dec <- filter(flights, month %in% c(11, 12))
```

###De-Morgan's law states that !(x|y) is the same as !x&!y
```{r message=FALSE, warning=FALSE}
filter(flights, !(arr_delay > 120 | dep_delay > 120))
filter(flights, arr_delay <= 120, dep_delay <= 120)
```

###missing values: filter excludes values that are FALSE and NA
```{r message=FALSE, warning=FALSE}
df <- tibble(x = c(1, NA, 3))
filter(df, x > 1)
```

###In case we want to exclude FALSE but retain NA, we write it as below
```{r message=FALSE, warning=FALSE}
filter(df, is.na(x) | x > 1)
```

###Arrange is used to order the data by columns; desc() is added to sort by descending order
```{r message=FALSE, warning=FALSE}
arrange(flights, year, month, day)
arrange(flights, desc(dep_delay))
```

###missing values are placed at the end when sorting regardless of ascending or descending
```{r message=FALSE, warning=FALSE}
df <- tibble(x = c(5, 2, NA))
arrange(df, x)
arrange(df, desc(x))
```

###Select() used to select specific columns; use 'a:b' operator to select all columns from a to b; use -() to exclude the columns in the parantheses; 
```{r message=FALSE, warning=FALSE}
select(flights, year, month, day)
select(flights, year:day)
select(flights, -(year:day))
```

###Rename used to rename column
```{r message=FALSE, warning=FALSE}
rename(flights, tail_num = tailnum)
```

###mutate() can be used to generate new columns after using select() to first select the columns we need
```{r message=FALSE, warning=FALSE}
flights_sml <- select(flights, 
  year:day, 
  ends_with("delay"), 
  distance, 
  air_time
)
mutate(flights_sml,
  gain = dep_delay - arr_delay,
  speed = distance / air_time * 60
)
```

###in mutate, we can create new columns referring to the new columns created right beforethe current one
```{r message=FALSE, warning=FALSE}
mutate(flights_sml,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)
```

###we use transmute() if the intention is to keep only the newly created columns and not all the old columns
```{r message=FALSE, warning=FALSE}
transmute(flights,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)
```

###%/% is used for integer division and %% is the reminder
```{r message=FALSE, warning=FALSE}
transmute(flights,
  dep_time,
  hour = dep_time %/% 100,
  minute = dep_time %% 100
)
```

###Offset functions: lead() moves the data forward one step, lag() lags the data by one step
```{r message=FALSE, warning=FALSE}
(x <- 1:10)
lag(x)
lead(x)
```

###cumsum can be used to calculate cumulative sum uptil the current datapoint; cummean is used to calculate cumulative mean uptil the current datapoint
```{r message=FALSE, warning=FALSE}
x
cumsum(x)
cummean(x)
```

###data can be ranked using min_rank() and min_rank(desc()) to rank the data
```{r message=FALSE, warning=FALSE}
y <- c(1, 2, 2, NA, 3, 4)
min_rank(y)
min_rank(desc(y))
```

###Other functions are summarized below
```{r message=FALSE, warning=FALSE}
row_number(y)
dense_rank(y)
percent_rank(y)
cume_dist(y)
```

###Summarise is used to calculate one metric using all the data
```{r message=FALSE, warning=FALSE}
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))
```

###Group_by groups the data by the chosen variables; summarise can folow this command to get the mean/median/variance etc. for all the variables we need
```{r message=FALSE, warning=FALSE}
by_day <- group_by(flights, year, month, day)
summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))
```

###Example case
```{r message=FALSE, warning=FALSE}
by_dest <- group_by(flights, dest)
##By destination, get number of entries, mean distance, mean delay time
delay <- summarise(by_dest,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE)
)
##filter the data with count>20 and destination not equal to HNL
delay <- filter(delay, count > 20, dest != "HNL")
```

###Plot the data of distance vs delay with the counts as size and plot a smooth line through the data
```{r message=FALSE, warning=FALSE}
ggplot(data = delay, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
```

###The filtering above can be carried out in a single step with pipes
```{r message=FALSE, warning=FALSE}
delays <- flights %>% 
  group_by(dest) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>% 
  filter(count > 20, dest != "HNL")
```


###In case of missing values, we have to explicitly call out na.rm=TRUE, else if there are any missing values, summary statistics will return NA
```{r message=FALSE, warning=FALSE}
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
##Above returns NA
##we have to add na.rm=TRUE
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay, na.rm = TRUE))
```

###In this context, we know that missing values represent canceled flights. so we can exclude them from the dataset to begin with
```{r message=FALSE, warning=FALSE}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))
```

###then we can summarise the data again as before
```{r message=FALSE, warning=FALSE}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
```

###Check average arrival delay by flight tail number
```{r message=FALSE, warning=FALSE}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay)
  )
```

###Plot the data as a frequency polygon
```{r message=FALSE, warning=FALSE}
ggplot(data = delays, mapping = aes(x = delay)) + 
  geom_freqpoly(binwidth = 10)
```

###As there are actually flights whose mean delay is over 5 hours, it makes sense to capture a count of the data for each tail number too and then plot a scatterplot
```{r message=FALSE, warning=FALSE}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay, na.rm = TRUE),
    n = n()
  )

ggplot(data = delays, mapping = aes(x = n, y = delay)) + 
  geom_point(alpha = 1/10)
```

###Next we filter out the data with very few datapoints so we can get a clearer view of the data
```{r message=FALSE, warning=FALSE}
delays %>% 
  filter(n > 25) %>% 
  ggplot(mapping = aes(x = n, y = delay)) + 
    geom_point(alpha = 1/10)
```

###Case to study average batting of batters
```{r message=FALSE, warning=FALSE}
batting <- as_tibble(Lahman::Batting)

##By Batter, calculate batting average and number of balls at bat
batters <- batting %>% 
  group_by(playerID) %>% 
  summarise(
    ba = sum(H, na.rm = TRUE) / sum(AB, na.rm = TRUE),
    ab = sum(AB, na.rm = TRUE)
  )

##Filter to those batters who saw more than 100 balls at bat and then plot batting average against balls at bat
batters %>% 
  filter(ab > 100) %>% 
  ggplot(mapping = aes(x = ab, y = ba)) +
    geom_point() + 
    geom_smooth(se = FALSE)
```

###Simple sorting of batting averages in descending order shows that they faced very few balls so they just got lucky
```{r message=FALSE, warning=FALSE}
batters %>% 
  arrange(desc(ba))
```

###get mean and mean for positive delays
```{r message=FALSE, warning=FALSE}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    avg_delay1 = mean(arr_delay),
    avg_delay2 = mean(arr_delay[arr_delay > 0]) # the average positive delay
  )
```

###calculate standard deviation of distance and sort by descending
```{r message=FALSE, warning=FALSE}
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(distance_sd = sd(distance)) %>% 
  arrange(desc(distance_sd))
```

###Get earliest and latest departure time
```{r message=FALSE, warning=FALSE}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    first = min(dep_time),
    last = max(dep_time)
  )
```

###first and last give us the first and last entries in the data; filter can be used to get the rows we want that meet the specific conditions
```{r message=FALSE, warning=FALSE}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    first_dep = first(dep_time), 
    last_dep = last(dep_time)
  )

not_cancelled %>% 
  group_by(year, month, day) %>% 
  mutate(r = min_rank(desc(dep_time))) %>% 
  filter(r %in% range(r))
```

###get distinct carriers
```{r message=FALSE, warning=FALSE}
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(carriers = n_distinct(carrier)) %>% 
  arrange(desc(carriers))
```

### for a simple count, we use count()
```{r message=FALSE, warning=FALSE}
not_cancelled %>% 
  count(dest)

##Count can be done with a weight varible to sum along that variable
not_cancelled %>% 
  count(tailnum, wt = distance)
```


```{r message=FALSE, warning=FALSE}
##Get number of flight leaving before 5AM
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(n_early = sum(dep_time < 500))

##Proportion of flights that are delayed by more than an hour
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(hour_perc = mean(arr_delay > 60))
```

###Use group_by to gradually roll-up the data by data, month, year
```{r message=FALSE, warning=FALSE}
daily <- group_by(flights, year, month, day)
(per_day   <- summarise(daily, flights = n()))

(per_month <- summarise(per_day, flights = sum(flights)))

(per_year  <- summarise(per_month, flights = sum(flights)))
```

###Ungroup data by using ungroup
```{r message=FALSE, warning=FALSE}
daily %>% 
  ungroup() %>%   
  summarise(flights = n())
```

###
```{r message=FALSE, warning=FALSE}
flights_sml %>% 
  group_by(year, month, day) %>%
  filter(rank(desc(arr_delay)) < 10)
```

###Get all groups that are bigger than a threshold
```{r message=FALSE, warning=FALSE}
popular_dests <- flights %>% 
  group_by(dest) %>% 
  filter(n() > 365)
popular_dests
```

###Standardize the arrival delays 
```{r message=FALSE, warning=FALSE}
popular_dests %>% 
  filter(arr_delay > 0) %>% 
  mutate(prop_delay = arr_delay / sum(arr_delay)) %>% 
  select(year:day, dest, arr_delay, prop_delay)
```