---
title: "Week 7"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Visualization 

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
```


###View the data in mpg; then create a scatter plot of the engine size and fuel efficiency
```{r message=FALSE, warning=FALSE}
mpg

ggplot(data=mpg)+geom_point(aes(x=displ, y=hwy))
  
```

###We can add the class of the car to the above plot by using color function in the aesthetics
```{r message=FALSE, warning=FALSE}
ggplot(data=mpg)+geom_point(aes(x=displ, y=hwy, color=class))
```

###We can also use size function instead of color to show each class of car as a size of the point in a scatter plot
```{r message=FALSE, warning=FALSE}
ggplot(data=mpg)+geom_point(aes(x=displ, y=hwy, size=class))
```

###We can also use the alpha function to map the class of a car to its transparency; instead, we can use the shape function to map the class of a car
```{r message=FALSE, warning=FALSE}
ggplot(data = mpg) + 
  geom_point(aes(x = displ, y = hwy, alpha = class))

ggplot(data = mpg) + 
  geom_point(aes(x = displ, y = hwy, shape = class))
```

###We can pick the color of the points for the scatter plot, it goes outside aesthetics
```{r message=FALSE, warning=FALSE}
ggplot(data = mpg) + 
  geom_point(aes(x = displ, y = hwy), color = "blue")
```

###Facets are used to split plots into multiple ones corresponding to each value of a categorical variable. Facet_wrap() is used to achieve this; class of the car is a categorical variable that can be faceted
```{r message=FALSE, warning=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

###We can use facet_wrap() to facet using two different categorical variables
```{r message=FALSE, warning=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)
```

###we can use geom type to decide the type of chart. Geom_point generates a scatter plot and geom_smooth plots a smooth curve that best fits the points
```{r message=FALSE, warning=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

###We can use linetype to plot different values of a categorical variable as a different type of smooothed line
```{r message=FALSE, warning=FALSE}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))

```

###We can overlap two plots to show both color and line type
```{r message=FALSE, warning=FALSE}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv, color = drv))+
  geom_point(mapping = aes(x = displ, y = hwy, color = drv))
```

###We can display multiple geom types in one plot by overlapping them
```{r message=FALSE, warning=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

###We can use color and smooth to show the points with different color types for class of cars and a smoothed geom through them; we can plot a smoothed line through a subset of the data, namely class = "subcompact"
```{r message=FALSE, warning=FALSE}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth()


ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)
```

### We can use bar plots to plot statistical distribution by categorical variables
```{r message=FALSE, warning=FALSE}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))

```

###we can use stats instead of geom_bar to get the same result, the count of each value of a categorical variable
```{r message=FALSE, warning=FALSE}
ggplot(data = diamonds) + 
  stat_count(mapping = aes(x = cut))
```

###The default type of the bar can be changed from count to actual value using "identity"
```{r message=FALSE, warning=FALSE}
demo <- tribble(
  ~cut,         ~freq,
  "Fair",       1610,
  "Good",       4906,
  "Very Good",  12082,
  "Premium",    13791,
  "Ideal",      21551
)

ggplot(data = demo) +
  geom_bar(mapping = aes(x = cut, y = freq), stat = "identity")
```

###We can use proportion instead of count to get proportion of population that a bar represents; We can use stat_summary() to get plots of statistical summaries by different variables
```{r message=FALSE, warning=FALSE}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))

ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )

```

###We can use color and fill to color the bar charts in different ways - border colors vs fill colors 
```{r message=FALSE, warning=FALSE}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, colour = cut))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = cut))
```

###We can use stacked bars by using two different variables in bar aesthetics
```{r message=FALSE, warning=FALSE}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity))
```

###We can  use alpha to manage transparency or fill = NA to not fill it with any color
```{r message=FALSE, warning=FALSE}
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + 
  geom_bar(alpha = 1/5, position = "identity")
ggplot(data = diamonds, mapping = aes(x = cut, colour = clarity)) + 
  geom_bar(fill = NA, position = "identity")
```

###position = fill can be use to create stacked bars that are off the same height for easier visual inspection
```{r message=FALSE, warning=FALSE}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")
```

###position = dodge can be used to stack bars next to each other instead of on top of each other
```{r message=FALSE, warning=FALSE}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
```

###we can use jitter to address the problem of overplotting by adding some random noise to each point
```{r message=FALSE, warning=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), position = "jitter")
```

###boxplot can be used to plot statistical distribution of data and can be shown either horizontally or vertically
```{r message=FALSE, warning=FALSE}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot()
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot() +
  coord_flip()
```

###coord_quickmap can be used to set aspect ratios for maps
```{r message=FALSE, warning=FALSE}
nz <- map_data("nz")

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black")

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black") +
  coord_quickmap()
```

###coord_polar can be used to plot polar coordinates
```{r message=FALSE, warning=FALSE}
bar <- ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = cut, fill = cut), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

bar + coord_flip()
bar + coord_polar()
```

