---
title: "Week 9"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Plotting Systems

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(lattice)
library(datasets)
```


###We can add a smooth line to highlight trends on plots
```{r message=FALSE, warning=FALSE}
data(airquality)

with(airquality, {plot(Temp, Ozone)+lines(loess.smooth(Temp, Ozone))})
  
```

###We can add annotations to plots such as a title
```{r message=FALSE, warning=FALSE}
  data(cars)
 with(cars, plot(speed, dist))
 title("Speed vs. Stopping distance")
```

###Lattice plots can be used to evaluate how one variable varies with another at different levels of a third variable. For example, it can be used to see how life expectancy and income vary by different regions
```{r message=FALSE, warning=FALSE}
state <- data.frame(state.x77, region = state.region)
xyplot(Life.Exp ~ Income | region, data = state, layout = c(4, 1))
```

###scatter plot can also be drawn with qplot
```{r message=FALSE, warning=FALSE}
data(mpg)
qplot(displ, hwy, data = mpg)
```

##Chapter 8: Graphic Devices

###One way to use plots is to save them in a file
```{r message=FALSE, warning=FALSE}
## Open a PDF file
pdf(file = "myplot.pdf")  
## Plot something that will be used to save to a file
with(faithful, plot(eruptions, waiting))  
## Add titles and other annotations as needed
title(main = "Old Faithful Geyser data")
## Close the PDF file you just opened and wrote the plot to
dev.off() 
```

###We can also copy plots we created to image files
```{r message=FALSE, warning=FALSE}
with(faithful, plot(eruptions, waiting)) 
title(main = "Old Faithful Geyser data") 

##Copy to an image file
dev.copy(png, file = "geyserplot.png") 
##Close the file that you opened to save to file
dev.off()  
```
